Gene expression

### 2022-11-30 code/gene_expression/Gene expression.Rmd
Final commit for submission of project

@@

## Statistical analysis of dCT levels 
```{r}
anova <- delta_gene_expr %>% 
  group_by(genes) %>% 
  anova_test(dct ~ groups) %>% 
  ungroup()

significant <- delta_gene_expr %>% 
  group_by(genes) %>% 
  do(lm(dct ~ groups,.) %>%  tidy) %>% 
  filter(term != "(Intercept)") %>% 
  filter(p.value <= 0.05)
```

@@

```{r}
delta_gene_expr %>% 
  group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(`2^-(ddct)`)) %>% 
  ggplot(
    aes(x = groups,
        y = mean_fold_change,
        fill = groups)
  ) + 
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )

delta_gene_expr %>% 
  group_by(genes, groups) %>% 
  ggplot(
    aes(x = groups,
        y = `2^-(ddct)`,
        fill = groups)
  ) + 
  geom_boxplot() +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )

@@

```{r}
my_comparisons <- list(c("VANCO-FVF", "VANCO-SM"), c("VANCO-BF", "VANCO-FVF"))
```

>>>>

library(g)
library(facetscales)
scales_y <- list(
  `4` = scale_y_continuous(limits = c(5, 25), breaks = seq(5, 25, 5)),
  `f` = scale_y_continuous(limits = c(0, 40), breaks = seq(0, 40, 10)),
  `r` = scale_y_continuous(limits = c(10, 20), breaks = seq(10, 20, 2))
)
ggplot(mpg, aes(displ, cty)) + 
  geom_point() + 
  facet_grid_sc(rows = vars(drv), scales = list(y = scales_y))

@@

## Statistical analysis of dCT levels 
```{r}
significant <- delta_gene_expr %>% 
  group_by(genes) %>% 
  do(lm(dct ~ groups,.) %>%  tidy) %>% 
  filter(term != "(Intercept)") %>% 
  filter(p.value <= 0.05)

anova <- delta_gene_expr %>% 
  group_by(genes) %>% 
  rstatix::anova_test(dct ~ groups)

tukey <- delta_gene_expr %>% 
  group_by(genes) %>% 
  rstatix::tukey_hsd(dct ~ groups) %>% 
  filter(p.adj < 0.05) %>% 
  flextable::as_flextable()

delta_gene_expr$group

@@

```{r}
delta_gene_expr %>% 
  group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(`2^-(ddct)`)) %>% 
  ggplot(
    aes(x = groups,
        y = mean_fold_change,
        fill = groups)
  ) + 
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )
```

## decreased_fold_change
```{r decreased-fold-change}
decreased_fold_change <- delta_gene_expr %>% 
  filter(genes %in% tukey$genes) %>% 
  filter(!genes %in% c("IL1b","REG3G")) %>% 
  mutate(across(genes, factor, levels=c("TGFb","cldn15","TJP1","VEGFa","Ocld1"))) %>%
  group_by(genes, groups) %>% 
  ggplot(
    aes(x = groups,
        y = `2^-(ddct)`,
        fill = groups)
  ) + 
  geom_boxplot(outlier.fill = NA,
               outlier.shape = NA) +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
   stat_compare_means(comparisons = my_comparisons, method = "wilcox.test", label.y = c(1.2,1.4)) +
  labs(
       y = "Fold Δ",
       x = "",
       fill = "Groups") +
  theme_bw()+
  theme(                   
    #axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.position = "none") + 
    scale_fill_brewer(palette="Set1")
  decreased_fold_change

@@

## Increased fold change
```{r increased-fold-change}
increased_fold_change <- delta_gene_expr %>% 
  #filter(genes %in% tukey$genes) %>% 
  filter(genes %in% c("IL1b","REG3G")) %>% 
 # mutate(across(genes, factor, levels=c("TGFb","cldn15","TJP1","VEGFa","Ocld1"))) %>%
  group_by(genes, groups) %>% 
  ggplot(
    aes(x = groups,
        y = `2^-(ddct)`,
        fill = groups)
  ) + 
  geom_boxplot(outlier.fill = NA,
               outlier.shape = NA) +
  facet_wrap(~genes,
             ncol = 2,
             scales = "free") +
   stat_compare_means(comparisons = my_comparisons, method = "wilcox.test", label.y.npc = c(0.5)) +
  labs(
       y = "Fold Δ",
       x = "",
       fill = "Groups") +
  theme_bw()+
  theme(                   
    #axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.position = "none") + 
    scale_fill_brewer(palette="Set1")
  increased_fold_change

```{r}
my_comparisons <- list(c("VANCO-BF", "VANCO-FVF"), c("VANCO-SM", "VANCO-FVF"))
```
